On the means, costs, and system-level impacts of 24/7 carbon-free energy procurement

### 2024-06-24 manuscript/manuscript.tex
revision: move sections into master tex

\setlength{\abovecaptionskip}{1pt} 	% Adjust the space above the caption
\graphicspath{ 						% Allows for multiple directories
	{},
	}

@@

	\author[tubaddress]{Iegor Riepin\fnref{fn1}}
	\ead{iegor.riepin@tu-berlin.de}

	\author[tubaddress]{Tom Brown}
	\ead{t.brown@tu-berlin.de}

	\address[tubaddress]{Department of Digital Transformation in Energy Systems, TU Berlin, Germany}

    \fntext[fn1]{Corresponding author: Iegor Riepin, Technical University of Berlin, Einsteinufer 25, 10587 Berlin, Germany}

	\begin{abstract}
		% The 5 things a good abstract needs:
		% 1. Introduce the topic: what we know
		% 2. State the unknown: what we don't know
		% 3. Outline the methods: how we answer the questions
		% 4. Preview the findings: what we learned
		% 5. Tell us what your work teaches us: what we conclude and why this is important
		% aim ~200 words | 225 words used
		A growing number of public and private energy buyers are interested in 24/7 carbon-free energy (CFE) procurement, which means that every kilowatt-hour of electricity consumption is met by carbon-free sources at all times.
		It has the potential to overcome the limitations of established procurement schemes, such as the temporal mismatch between clean electricity supply and buyers' demand that is inherent to \enquote{volumetric} matching.
		Yet it is unclear how 24/7 CFE procurement affects the rest of the power system, and whether this effect is consistent across regional contexts and different levels of system cleanness.
		We use a mathematical model to systematically examine different designs, optimal procurement strategies, costs, and impacts of the 24/7 CFE matching, both for participating buyers and for regions where voluntary procurement occurs.
		We examine mechanisms driving system-level emissions reduction and how they vary across regions and over time.
		Our results indicate that clean energy procurement commitments have consistent beneficial effects on participants and the electricity system.
		Even as grids become cleaner over time, the hourly matching strategy contributes significantly to system-level emissions reduction.
		In addition, voluntary commitments to 24/7 CFE have a further transformative effect on electricity systems through accelerated innovation and early deployment of advanced energy technologies.

@@

\tableofcontents

\section{Introduction}
\label{sec:intro}
% Introduction Structure

@@

Climate change is driving a global effort to rapidly decarbonise electricity systems across the globe.
The European Union (\gls{eu}) has committed to achieving net-zero emissions by 2050, and the European Commission has adopted the European Climate Law with a target of reducing net greenhouse gas emissions by at least 55\% by 2030, compared to 1990 levels \cite{EuropeanClimateLaw2020}.
National climate initiatives are gaining ground across the globe, including the United States with the Biden administration's goal of reducing emissions by 50-52\% by 2030, compared to 2005 levels \cite{BidenClimatePlan2021}, and China's commitment to have CO$_2$ emissions peak before 2030 and achieve carbon neutrality before 2060 \cite{ChinaNetZero-IEA, liuEconomicCostEnergy2023}.

@@

Google initiated the 24/7 movement in 2020 by announcing its commitment to cover the demand at all its data centres and campuses worldwide using the 24/7 CFE approach by 2030 \cite{google-247by2030}.
Shortly after, Google published a policy roadmap on achieving the goal \cite{google-PolicyRoadmap} and methodologies paper for how to measure the progress \cite{google-methodologies}.
Also, Microsoft has committed to the 100/100/0 by 2030 goal (i.e. \enquote{100\% of electricity consumption, 100\% of the time, matched by zero carbon energy purchases} \cite{Microsoft-vision}.
The list of entities that have committed to 24/7 CFE procurement includes also utilities, such as Peninsula Clean Energy (a community choice aggregator in California) committed to 24/7~CFE by 2025 \cite{peninsula-OurPathto247}, and even cities like Des Moines, Iowa U.S., with a commitment by 2035 \cite{iowaenvcouncil-247}. In 2021, an international group of energy suppliers, electricity buyers, and \gls{ngo}s launched the 24/7 Carbon-free Energy Compact \cite{gocarbonfree247}.
With over 135 signatories, the group aims to \enquote{develop and scale high-impact technologies, energy policies, procurement practices, and solutions to make 24/7 Carbon-Free Energy achievable for all}.

@@

\textbf{Optimisation model --} The simulations were carried out with the modified version of the open European power system model PyPSA-Eur, built using the open framework PyPSA \cite{brownPyPSAPythonPower2018}.
The objective of the model is to minimise the annualised system-wide costs of meeting energy demand within the modelled system, while adhering to all relevant engineering, reliability, and policy constraints.
We use a brownfield investment approach, which means that we keep the existing power generation and storage assets in the model and only optimise the capacity expansion of new assets.
The optimal solution of the model includes the location and capacity of new power generation and storage assets, as well as the hourly dispatch of all existing and new assets, including the linear optimal power flow on the transmission network for the planning year.
This is standard for the PyPSA-Eur model \cite{horschPyPSAEurOpenOptimisation2018} and the long-term energy planning models of this class \cite{jenkinsGenX2022, howellsOSeMOSYSOpenSource2011}.

\textbf{Clean energy procurement model --} We introduce new model components (parameters, variables, constraints) into the PyPSA-Eur model to represent voluntary clean energy procurement.
By incorporating these into the mathematical problem, we model a situation where a fraction of electricity demand in a particular bidding zone is voluntarily committed to procuring clean energy with a desired strategy.
The fraction is set at 10\% of the total \gls{ci} electricity demand in a particular bidding zone, hereinafter referred to as \enquote{participating consumers}.
The model co-optimises procurement and operational decisions of the participating consumers to meet their electricity demand in accordance with the procurement strategy they have chosen.

@@

% On bilinear term due to grid CFE factor
The CFE factor of the regional grid ($CFE_t$) can be seen as the percentage of clean electricity in each MWh of imported electricity to supply demand of participating consumers in a given hour.
To compute $CFE_t$, we consider both the hourly electricity mix in the local bidding zone and emission intensity of imported electricity from the neighbouring zones.
The methodology to calculate the CFE factor of the regional grid is described in detail in the prior work of the authors \cite{riepin-zenodo-systemlevel247}.

@@

\textbf{Model scope --} Geographically, the model encompasses the entire \gls{entsoe} area, which covers the entire European electricity system.\footnote{Islands that are not connected to the main European system, such as Malta, Crete and Cyprus, are excluded from the model.}
In this study, electricity demand, supply and power transmission infrastructure (alternating current lines at and above 220~kV voltage level and all high voltage direct current lines) are clustered to 37 individual bidding zones \cite{PyPSAEur-docs-spatialresolution}.
Thus, each zone represents a country; some countries straddling different synchronous areas are split into individual bidding zones, such as DK1 (West) and DK2 (East).

Participating consumers can be located in either of the four selected zones: Ireland, Denmark (zone DK1), Germany, or Poland.
The comparison of results across these zones allows us to generalize the implications of voluntary clean energy procurement, taking into account factors such as electricity demand, weather, availability of renewable resources, existing electricity generation capacity mix, and national energy policies, among other factors.

@@

\textbf{Technology assumptions --} Cost and other assumptions for energy technologies available for participating consumers were collected primarily from the Danish Energy Agency \cite{DEA-technologydata} for the respective years and are provided in [\cref{tab:tech_costs}].
For the \gls{ldes}, we assume an underground hydrogen storage in salt caverns with an electolyzer and a fuel cell for hydrogen conversion.
Data for advanced clean firm technologies is less reliable due to technological uncertainty and lack of commercial experience; therefore, we use our own indicative assumptions.

@@

Throughout the analysis, we examine how voluntary clean electricity procurement strategies affect both buyers and the electricity system as a whole.
From the \gls{ci} buyers' perspective, we examine optimal procurement strategies (i.e., a mix of locally procured energy generation and storage assets and electricity imports), the average cost of procured electricity and the carbon emissions associated with electricity usage, also known as Scope 2 emissions \cite{GHGProtocolScope2}.
From a system perspective, we analyze how voluntary clean energy procurement impacts CO$_2$ emissions in the region and the \enquote{demand pull} for advanced energy technologies.

@@

\begin{res}
    The disparity in decarbonisation outcomes between 24/7 CFE hourly and 100\% annual renewable matching policies becomes increasingly pronounced as local grids transition to cleaner states. This phenomenon underscores the effectiveness of 24/7 CFE procurement in supporting system decarbonisation in the context of evolving national energy and climate policies.
\end{res}

@@

% Limitations: a point on DSM
Fourth, participating consumers have an inflexible demand, i.e., no demand-side management is considered.
In reality, many \gls{ci} consumers have some degree of demand flexibility, which they can use to adjust their consumption as needed based on grid signals, such as the wholesale market price and the average emission rate of the local electricity mix.
Moreover, some consumers, such as data centers and cloud computing providers, can shift their load also in space, i.e., to shift computing
jobs and associated power loads across data center locations \cite{rosskoningsteinWeNowMore2021}.
In the recent work, Riepin~\&~Brown (2023) explore how space-time load-shifting flexibility can be used to meet high 24/7 CFE targets, as well as what potential benefits it may offer to participating energy buyers and to the rest of the energy system \cite{riepin-zenodo-spacetime247CFE}.

@@

In this work, we systematically examined the means, costs, and the system-level impacts of the 24/7 carbon-free energy matching.
We find that 24/7 matching is a viable and effective strategy for electricity buyers aiming to eliminate their own carbon footprint and contribute to wider system decarbonisation.
Voluntary commitments to the hourly matching strategy have a further transformative effect on electricity systems through accelerated innovation and early deployment of advanced energy technologies.
The evidence holds across geographical locations and is robust to sensitivity analyses.
Even as grids become cleaner over time, the 24/7 matching strategy retains its significance in reducing system-level emissions, which is essential in light of evolving national energy and climate policies.

@@

%S2: Background system assumptions

@@

>>>>

\setlength{\abovecaptionskip}{1pt} 	% Adjust the space above the caption
\graphicspath{ 						% Allows for multiple directories
	{../results/paper/},
	}

@@

	\author[tubaddress]{Iegor Riepin\fnref{fn1}}
	\ead{iegor.riepin@tu-berlin.de}
	\author[tubaddress]{Tom Brown}
	\ead{t.brown@tu-berlin.de}

	\address[tubaddress]{Department of Digital Transformation in Energy Systems, TU Berlin, Germany}
	\fntext[fn1]{Corresponding author: Iegor Riepin, Technical University of Berlin, Einsteinufer 25, 10587 Berlin, Germany}

\begin{abstract}
% The 5 things a good abstract needs:
% 1. Introduce the topic: what we know
% 2. State the unknown: what we don't know
% 3. Outline the methods: how we answer the questions
% 4. Preview the findings: what we learned
% 5. Tell us what your work teaches us: what we conclude and why this is important
% aim ~200 words | working paper: 205 words | revision: 230 words

A growing number of public and private energy buyers are interested in 24/7 carbon-free energy (CFE) procurement, which means that every kilowatt-hour of electricity consumption is met by carbon-free sources at all times.
It has the potential to overcome the limitations of established procurement schemes, such as the temporal mismatch between clean electricity supply and buyers' demand that is inherent to \enquote{volumetric} matching.
Yet it is unclear how 24/7 CFE procurement affects the rest of the power system, and whether this effect is consistent across regional contexts and different levels of system cleanness.
We use a mathematical model to systematically examine different designs, optimal procurement strategies, costs, and impacts of the 24/7 CFE matching, both for participating buyers and for regions where voluntary procurement occurs.
We examine mechanisms driving system-level emissions reduction and how they vary across regions and over time.
Our results indicate that clean energy procurement commitments have consistent beneficial effects on participants and the electricity system.
Utilizing an hourly matching strategy, energy buyers can eliminate all carbon emissions linked to their electricity consumption, while also reducing system-level emissions by up to 572~kgCO$_2$/a per MWh of participating load.
Even as grids become cleaner over time, the hourly matching strategy contributes significantly to system-level emissions reduction.
In addition, voluntary commitments to 24/7 CFE have a further transformative effect on electricity systems through accelerated innovation and early deployment of advanced energy technologies.

@@

% \tableofcontents

\section{Introduction}
\label{sec:intro}

@@

Climate change is driving a global effort to rapidly decarbonise electricity systems across the globe.
The European Union (\gls{eu}) has committed to achieving net-zero emissions by 2050, and the European Commission has adopted the European Climate Law with a target of reducing net greenhouse gas emissions by at least 55\% by 2030, compared to 1990 levels \cite{EuropeanClimateLaw2020}.
National climate initiatives are gaining ground across the globe, including the United States with the Biden administration's goal of reducing emissions by 50-52\% by 2030, compared to 2005 levels \cite{BidenClimatePlan2021}, and China's commitment to have CO$_2$ emissions peak before 2030 and achieve carbon neutrality before 2060 \cite{ChinaNetZero-IEA}.

@@

Google initiated the 24/7 CFE movement in 2020 by announcing its commitment to cover the demand at all its data centres and campuses worldwide using the 24/7 CFE approach by 2030 \cite{google-247by2030}.
Shortly after, Google published a policy roadmap on achieving the goal \cite{google-PolicyRoadmap}.	Shortly after, Google published a policy roadmap on achieving the goal \cite{google-PolicyRoadmap} and methodologies paper for how to measure the progress \cite{google-methodologies}.
Also, Microsoft has committed to the 100/100/0 by 2030 goal (i.e. \enquote{100\% of electricity consumption, 100\% of the time, matched by zero carbon energy purchases} \cite{Microsoft-vision}.
The list of entities that have committed to 24/7 CFE procurement includes also utilities, such as Peninsula Clean Energy (a community choice aggregator in California) committed to 24/7~CFE by 2025 \cite{peninsula-OurPathto247}, and even cities like Des Moines, Iowa U.S., with a commitment by 2035 \cite{iowaenvcouncil-247}. In 2021, an international group of energy suppliers, electricity buyers, and \gls{ngo}s launched the 24/7 Carbon-free Energy Compact \cite{gocarbonfree247}.
With over 135 signatories, the group aims to \enquote{develop and scale high-impact technologies, energy policies, procurement practices, and solutions to make 24/7 Carbon-Free Energy achievable for all}.

@@

\textbf{Optimisation model --} The simulations were carried out with the modified version of the open European power system model PyPSA-Eur, built using the open framework PyPSA \cite{brownPyPSAPythonPower2018}.
The objective of the model is to minimise the annualised system-wide costs of meeting energy demand within the modelled system, while adhering to all relevant engineering, reliability, and policy constraints.
We use a brownfield investment approach, which means that we keep the existing power generation and storage assets in the model and only optimise the capacity expansion of new assets.
The optimal solution of the model includes the location and capacity of new power generation and storage assets, as well as the hourly dispatch of all existing and new assets, including the linear optimal power flow on the transmission network for the planning year.
This is standard for the PyPSA-Eur model \cite{horschPyPSAEurOpenOptimisation2018} and the long-term energy planning models of this class \cite{jenkinsGenX2022, howellsOSeMOSYSOpenSource2011}.
The mathematical formulation of the model is available in PyPSA documentation \cite{pypsadocs}.
The code to reproduce the optimisation problem behind this analysis is available at GitHub \cite{github-247CFEpaper}.

\textbf{Clean energy procurement model --} We introduce new model components (parameters, variables, constraints) into the PyPSA-Eur model to represent voluntary clean energy procurement.
By incorporating these into the mathematical problem, we model a situation where a fraction of electricity demand in a particular bidding zone is voluntarily committed to procuring clean energy with a desired strategy.
The fraction is set at 10\% of the total \gls{ci} electricity demand in a particular bidding zone, representing a hypotetical group of electricity consumers (e.g. datacentres, industrial facilities, or municipalities) committed to clean energy procurement, hereinafter referred to as \enquote{participating consumers}.
The model seeks to find cost-optimal procurement strategy and operational decisions for the participating consumers to meet their electricity demand in accordance with the chosen clean electricity matching strategy.

@@

% On bilinear term due to grid CFE factor
The CFE factor of the regional grid ($CFE_t$) can be seen as the percentage of clean electricity in each MWh of imported electricity to supply demand of participating consumers in a given hour.
To compute $CFE_t$, we consider both the hourly electricity mix in the local bidding zone and emission intensity of imported electricity from the neighbouring zones.
The methodology to calculate the CFE factor of the grid is described in the prior work of the authors \cite{riepin-zenodo-systemlevel247}.

@@

\textbf{Model scope --} Geographically, the model encompasses the entire \gls{entsoe} area, which covers the entire European electricity system.\footnote{Islands that are not connected to the main European system, such as Malta, Crete and Cyprus, are excluded from the model.}
In this study, electricity demand, supply and power transmission infrastructure (alternating current lines at and above 220~kV voltage level and all high voltage direct current lines) are clustered to 37 individual bidding zones \cite{PyPSAEur-docs}.
Thus, each zone represents a country; some countries straddling different synchronous areas are split into individual bidding zones, such as DK1 (West) and DK2 (East).

Participating consumers can be located in either of the four selected zones: Ireland, Denmark (zone DK1), Germany, or Poland.
The comparison of results across these zones allows us to generalize the implications of voluntary clean energy procurement, taking into account individual characteristics of each zone, such as electricity demand, weather, availability of renewable resources, existing electricity generation capacity mix, and national energy policies, among other factors.

@@

\textbf{Technology assumptions --} Cost and other assumptions for energy technologies available for participating consumers were collected primarily from the Danish Energy Agency \cite{DEA-technologydata} for the respective years and are provided in [\cref{tab:tech_costs}].
For the \gls{ldes}, we assume an underground hydrogen storage in salt caverns with an electolyzer and a fuel cell for hydrogen conversion.
Data for advanced clean firm technologies is less reliable due to technological uncertainty and lack of commercial experience; therefore, we use own indicative assumptions.

@@

Throughout the analysis, we examine how voluntary clean electricity procurement strategies affect both buyers and the electricity system as a whole.
From the \gls{ci} buyers' perspective, we examine optimal procurement strategies for both annual and hourly matching strategies (i.e., a cost-optimal mix of locally procured energy generation, storage assets and electricity imports), the average cost of procured electricity and the carbon emissions associated with electricity usage, also known as Scope 2 emissions \cite{GHGProtocolScope2}.
From a system perspective, we analyze how voluntary clean energy procurement impacts CO$_2$ emissions in the background grid and the \enquote{demand pull} for advanced energy technologies.

@@

\begin{res}
    The disparity in decarbonisation outcomes between 24/7 CFE hourly and 100\% annual renewable matching policies becomes increasingly pronounced as local grids transition to cleaner states over time.
\end{res}

@@

% Limitations: a point on DSM
Fourth, participating consumers have an inflexible demand, i.e., no demand-side management is considered.
In reality, many \gls{ci} consumers have some degree of demand flexibility, which they can use to adjust their consumption as needed based on grid signals, such as the wholesale market price and the average emission rate of the local electricity mix.
Moreover, some consumers, such as data centers and cloud computing providers, can shift their load also in space, i.e., to shift computing jobs and associated power loads across data center locations \cite{koningsteinWeNowMore2021}.
The ability to shift power loads in time and space can play a significant role in achieving hourly carbon-free energy matching by 2030, as more applications and engineering solutions become available \cite{radovanovicCarbonAwareComputingDatacenters2021,radovanovicIEEE2023,koningsteinWeNowMore2021}.
In the recent work, Riepin~\&~Brown (2023) explore how space-time load-shifting flexibility can be used to meet high 24/7 CFE targets. \cite{riepin-zenodo-spacetime247CFE}.

@@

In this work, we systematically examined the means, costs, and the system-level impacts of the 24/7 carbon-free energy matching.
We find that 24/7 matching is a viable and effective strategy for clean electricity buyers aiming to eliminate their own carbon footprint and contribute to wider system decarbonisation.
The evidence holds across different geographical locations.
Even as grids become cleaner over time, the 24/7 matching strategy retains its significance in reducing system-level emissions, which is essential in light of evolving national energy and climate policies.
Voluntary commitments to the hourly matching strategy have a further transformative effect on electricity systems by creating an early market for advanced energy technologies, such as long-duration energy storage and clean firm power generators.

@@

@@

% \addcontentsline{toc}{section}{Supplementary References}
% \renewcommand{\ttdefault}{\sfdefault}
% \bibliographyS{paper_247CFE}
